A Loop Quantm Gravity Inspired Action for the Bosonic String and Emergent Dimensions at Large Scales

### 2024-12-02 lqg-strings.tex
Typo in title fixed
sketched beggining of solving gauged sigma model and made some corrections to Introduction in main text

\title{A Loop Quantm Gravity Inspired Action for the Bosonic String and a Emergent Dimensions at Large Scales}

@@

In the decades prior to the advent of string theory a great deal of effort was put into attempting to quantize gravity via both the Lagrangian and Hamiltonian approaches. The Hamiltonian approach involves, as stated earlier, a choice of the foliation of the background geometry into a collection of spacelike surfaces $\Sigma_t$. Central to this approach was the ADM (Arnowitt-Deser-Misner) formalism which allows one to construct a Hamiltonian $H_{GR}$ for general relativity starting from the Einstein-Hilbert action $S_{EH}$. The resulting phase space is coordinatized by a configuration variable which is the 3-metric $h_{ab}$ of the ``leaves'' $\Sigma_t$ of the foliation and a momentum variable which is a function of the extrinsic curvature $k_{ab}$ of the leaves\footnote{Our notational convention is the following. Lowercase Greek letters from the middle of the alphabet $\mu, \nu, \ldots$ are spacetime indices which run from $(0 \ldots 3)$, whereas lowercase Roman letters $a,b,\ldots$ are spatial indices for quantities which live solely on the surfaces $\Sigma_t$ and take values in $(1,2,3)$}. The gravity Hamiltonian turns out to be a sum of two constraints known as the diffeomorphism (or ``momentum'') constraint $H_{diff}$ and the Hamiltonian constraint $H_{ham}$:
\begin{equation}
	H_{GR} = H_{diff} + H_{ham},
\end{equation}
and these are, in turn, functionals on the phase space of general relativity written in terms of $h_{ab}$ and $k_{ab}$. The idea then is that physical states of the theory $\ket{\Psi_{phys}} $, which can be written as functionals $\Psi[h_{ab}]$ of the 3-metric $h_{ab}$ , must be annihilated by these constraints:
\begin{equation}
	H_{GR} \ket{\Psi_{phys}} \equiv 0
\end{equation}
While this procedure is straightforward in principle, in practice it was impossible to implement in the quantum theory due to the highly complicated non-polynomial dependence of the diffeomorphism and Hamiltonian constraints on the configuration and momentum variables. This unfortunate state of affairs persisted until the late 1980s when Abhay Ashtekar \cite{Ashketar1986New,Ashtekar1987New} recast general relativity as a theory of a densitized Minkowski tetrad
\begin{equation} \label{eqn:densitized-tetrad}
	\widetilde{E}^{\mu} _I = \frac{1}{\sqrt{- g}} e^{\mu} _I
\end{equation}
and the self-dual part of the $\sltwoc$ spin-connection
\begin{equation} \label{eqn:self-dual-spin-connection}
	A_{\mu} ^{I J} = \omega ^{I J} _{\mu} - \frac{i}{2} \varepsilon ^{I J} _{\ \ K L} \omega ^{K L} _{\mu} .
\end{equation}
Ashtekar realized that the constraints of general relativity, when expressed in terms of these ``new variables'', simplified drastically and became polynomial functions of the tetrad and connection variables. It was quickly realized by various researchers that the resulting theory could be quantized using the same methods used to quantize Yang-Mills gauge theory \cite{Ashtekar1992Weaving,Ashtekar2004Background} and that the diffeomorphism constraint could be solved exactly in terms of so-called \emph{spin-network} states \cite{Rovelli1988Knot,Rovelli1990Loop,Ashtekar1992Gravitons}.

>>>>

\title{A Loop Quantm Gravity Inspired Action for the Bosonic String and Emergent Dimensions at Large Scales}

@@

In the decades prior to the advent of string theory a great deal of effort was put into attempting to quantize gravity via both the Lagrangian and Hamiltonian approaches. The Hamiltonian approach involves, as stated earlier, a choice of the foliation of the background geometry into a collection of spacelike surfaces $\Sigma_t$. Central to this approach was the ADM (Arnowitt-Deser-Misner) formalism which allows one to construct a Hamiltonian $H_{GR}$ for general relativity starting from the Einstein-Hilbert action $S_{EH}$. The resulting phase space is coordinatized by a configuration variable which is the 3-metric $h_{ab}$ of the ``leaves'' $\Sigma_t$ of the foliation and a momentum variable which is a function of the extrinsic curvature $k_{ab}$ of the leaves\footnote{Our notational convention is the following. Lowercase Greek letters from the middle of the alphabet $\mu, \nu, \ldots$ are spacetime indices which run from $(0 \ldots 3)$, whereas lowercase Roman letters $a,b,\ldots$ are spatial indices for quantities which live solely on the surfaces $\Sigma_t$ and take values in $(1,2,3)$}. The gravity Hamiltonian turns out to be a sum of two constraints known as the diffeomorphism (or ``momentum'') constraint $H_{diff}$ and the Hamiltonian constraint $H_{ham}$:
\begin{equation}
	H_{GR} = H_{diff} + H_{ham},
\end{equation}
and these are, in turn, functionals on the phase space of general relativity written in terms of $h_{ab}$ and $k_{ab}$. The idea then is that physical states of the theory $\ket{\Psi_{phys}} $, which can be written as functionals $\Psi[h_{ab}]$ of the 3-metric $h_{ab}$ , must be annihilated by these constraints:
\begin{equation}
	H_{GR} \ket{\Psi_{phys}} \equiv 0
\end{equation}
While this procedure is straightforward in principle, in practice it was impossible to implement in the quantum theory due to the highly complicated non-polynomial dependence of the diffeomorphism and Hamiltonian constraints on the configuration and momentum variables. This unfortunate state of affairs persisted until the late 1980s when Abhay Ashtekar \cite{Ashketar1986New,Ashtekar1987New} recast general relativity as a BF theory of a dualised $\mathfrak{sl} (2 , \C)$ connection (which becomes the self-dual part of said connection in the case of $\beta = i$)
\begin{equation} \label{eqn:self-dual-spin-connection}
	A_{\mu} ^{I J} = \omega ^{I J} _{\mu} - \frac{\beta}{2} \varepsilon ^{I J} _{\ \ K L} \omega ^{K L} _{\mu}
\end{equation}
and its conjugate momenta
\begin{equation} \label{eqn:densitized-tetrad}
	B^{I J} _{\mu \nu} = \left( \delta ^{I} _{K} \delta ^{J} _{L} + \frac{1}{2 \beta} \varepsilon ^{I J} _{\ \ K L} \right) e^{K} _{[ \mu} e^{L} _{\nu ]} ,
\end{equation}
which when restricted to a 3-dimensional hypersurface where $* (e \wedge e) = 0$ reduces to
\begin{equation} \label{eqn:Ashtekar-connection-canonical}
	A^{i} _{a} = \omega ^{i} _{a} + \beta K^{i} _{a}
\end{equation}
\begin{equation} \label{eqn:Ashtekar-E-field-canonical}
	E^{a} _{i} = \sqrt{q} e^{a} _{i} = \frac{1}{2} \varepsilon ^{a b c} \varepsilon _{i j k} e^{j} _{b} e^{k} _{c} ,
\end{equation}
with $K^{i} _{a}$ being the triad version of the extrinsic curvature $K_{a b} = \frac{1}{2} \mathcal{L}_{n} q_{a b}$ and $q$ the 3-metric. Ashtekar realized that the constraints of general relativity, when expressed in terms of these ``new variables'', simplified drastically and became polynomial functions of the tetrad and connection variables. It was quickly realized by various researchers that the resulting theory could be quantized using the same methods used to quantize Yang-Mills gauge theory \cite{Ashtekar1992Weaving,Ashtekar2004Background} and that the diffeomorphism constraint could be solved exactly in terms of so-called \emph{spin-network} states \cite{Rovelli1988Knot,Rovelli1990Loop,Ashtekar1992Gravitons}.
